---
title: "StartCleaning"
author: "Nicholas Mobley"
date: "11/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(gridExtra)
```

Reading in the data
```{r}
kki.demo <- read.csv("C:\\Users\\nmobl\\Desktop\\School\\Fall2019\\STAT-S431\\S431Project\\Data\\KKI_demographicInfo.csv")
kki.social <- read.csv("C:\\Users\\nmobl\\Desktop\\School\\Fall2019\\STAT-S431\\S431Project\\Data\\KKI_SocialResponsivenessScaleQuestionnaire.csv")
kki.movement <- read.csv("C:\\Users\\nmobl\\Desktop\\School\\Fall2019\\STAT-S431\\S431Project\\Data\\KKI_movementAssessmentBatteryforChildren.csv")
```

```{r}
#Devloping a color pallete for the different groups.
OurColors <- brewer.pal(3,"Dark2")
names(OurColors) <- levels(kki$PrimaryDiagnosis)
colScale <- scale_color_manual(name = 'PrimaryDiagnosis',values = OurColors)
```





```{r}
#Fix odd ID col
colnames(kki.demo)[1] <- "ID"
colnames(kki.social)[1] <- "ID"
colnames(kki.movement)[1] <- "ID"
kki <- merge(kki.demo,kki.social)
kki <- merge(kki,kki.movement)
#Now have all the data in a huge data frame

#Removing N/A values. Mostly removing second visits.
kki <- na.omit(kki)

```

Exploring relationship between motor skills and social deficits
in children with ADHD and Autism

```{r}
#basic scatterplot
ggplot(data = kki, aes(x=mABC_TotalStandardScore, y =SRS_TotalRawScore)) + geom_point(alpha = .05) + geom_jitter() +geom_smooth()
```

```{r, warning = FALSE}
ggplot(data = kki, aes(x = mABC_TotalStandardScore, y = SRS_TotalRawScore,colour = PrimaryDiagnosis)) + geom_point(alpha = .3) + geom_jitter() + geom_smooth()
## Really one ADHD Shows any relationship
```

```{r}
##Another representation of previous plot
ggplot(data = kki, aes(x = mABC_TotalStandardScore, y = SRS_TotalRawScore,colour = PrimaryDiagnosis)) + geom_boxplot()
```

```{r}
Move.Age <- ggplot(data = kki, aes(x = mABC_AGE, y = mABC_TotalStandardScore, colour = PrimaryDiagnosis)) + geom_point(alpha = .3) + geom_jitter() + geom_smooth() + theme(legend.position = 'none')
#low correlation between age and motor skills in all 
# groups
```

```{r}
SRS.Age <- ggplot(data = kki, aes(x = SRS_AGE, y = SRS_TotalRawScore, colour = PrimaryDiagnosis)) + geom_point(alpha = .3) + geom_jitter() + geom_smooth() + theme(legend.position = 'none')
#again no real relationship
```

```{r}
grid.arrange(Move.Age,SRS.Age, nrow = 1)
#I don't know a good way of putting the gird together
# and the legend so somebody might wanna try that
```

```{r}
SRS.Manual <- ggplot(data = kki, aes(x = mABC_ManualDexterity.Component.StandardScore, y = SRS_TotalRawScore, colour = PrimaryDiagnosis)) + geom_point(alpha = .3) + geom_jitter() + geom_smooth()+ theme(legend.position = 'none')
```

```{r}
SRS.AimAndCatch <- ggplot(data = kki, aes(x = mABC_AimingAndCatching.Component.StandardScore, y = SRS_TotalRawScore, colour = PrimaryDiagnosis)) + geom_point(alpha = .3) + geom_jitter() + geom_smooth()+ theme(legend.position = 'none')
```

```{r}
#Really interesting! 
#The better the ADHD kids were at balance the better their social deficit was! Let's investigate the relationship between age and balance
SRS.Balance <- ggplot(data = kki, aes(x = mABC_Balance.Component.StandardScore, y = SRS_TotalRawScore, colour = PrimaryDiagnosis)) + geom_point(alpha = .3) + geom_jitter() + geom_smooth() + theme(legend.position = 'none')
```

```{r}
grid.arrange(SRS.Manual,SRS.AimAndCatch,SRS.Balance, nrow = 2)
```





```{r}
Balance.Age <- ggplot(data = kki, aes(y = mABC_Balance.Component.StandardScore, x = mABC_AGE, colour = PrimaryDiagnosis)) + geom_point(alpha = .3) + geom_jitter() + geom_smooth() + theme(legend.position = 'none')
```

```{r}
grid.arrange(Balance.Age,SRS.Age, nrow = 1)
```


### So far the only meaningfull conclusion I have is that when a child with ADHD has really good balance they have less social deficit. Doesn't seem to have a correlation to the other two movement tests. We might want to consider having a focus on the balance test.

Below will be subdividing all the tests into the three seperate tests conducted


```{r}
SRS.Bal1 <- ggplot(data = kki, aes(y = SRS_TotalRawScore, x = mABC_Balance.1.StandardScore, colour = PrimaryDiagnosis)) + geom_point(alpha = .8) + geom_jitter() + geom_smooth() + theme(legend.position = 'none')
```

```{r}
SRS.Bal2 <- ggplot(data = kki, aes(y = SRS_TotalRawScore, x = mABC_Balance.2.StandardScore, colour = PrimaryDiagnosis)) + geom_point(alpha = .8) + geom_jitter() + geom_smooth() + theme(legend.position = 'none')
```

```{r}
SRS.Bal3 <- ggplot(data = kki, aes(y = SRS_TotalRawScore, x = mABC_Balance.3.StandardScore, colour = PrimaryDiagnosis)) + geom_point(alpha = .8) + geom_jitter() + geom_smooth() + theme(legend.position = 'none')
```

```{r}
grid.arrange(SRS.Bal1,SRS.Bal2,SRS.Bal3,nrow = 2)
```


This looks to tell me that some balance tests were easier than others with 3 being the easiest, 2 being the hardest, and 1 being more well rounded








