---
title: "S431 Project EDA- Alli"
author: "Allison McCarty"
date: "11/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####My goal is to determine the relationship between social deficits and motor skill for children with multiple diagnoses. To do this, I want to see if having multiple diagnoses correlates to a change in a change in motor skills (compared to only having one diagnosis)
```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(gridExtra)
library(ggthemes)
```

Reading in the data
```{r, message=FALSE}
kki.demo <- read.csv("KKI_demographicInfo.csv")
kki.social <- read.csv("KKI_SocialResponsivenessScaleQuestionnaire.csv")
kki.movement <- read.csv("KKI_movementAssessmentBatteryforChildren.csv")
```

```{r, message=FALSE, warning=FALSE}
#Fix odd ID col
colnames(kki.demo)[1] <- "ID"
colnames(kki.social)[1] <- "ID"
colnames(kki.movement)[1] <- "ID"
kki <- merge(kki.demo,kki.social)
kki <- merge(kki,kki.movement)
#Now have all the data in a huge data frame
#Removing N/A values. Mostly removing second visits.
kki <- na.omit(kki)
kki2 <- filter(kki,PrimaryDiagnosis != "None",SecondaryDiagnosis ==c("Yes", "No"))
multdia <- filter(kki, PrimaryDiagnosis != "None", SecondaryDiagnosis =="Yes")
singdia <- filter(kki, PrimaryDiagnosis != "None" , SecondaryDiagnosis == "No")
nodia <- filter(kki, PrimaryDiagnosis =="None" , SecondaryDiagnosis =="No")

```


For this graph, we filtered out typically developing children to look at the how having additional diagnoses affects the relationship between motor skills and social development. 

```{r}
ggplot(data = kki2, aes(x = mABC_TotalStandardScore, y = SRS_TotalRawScore,colour = SecondaryDiagnosis)) + geom_point() + geom_jitter() + geom_smooth(se=FALSE)
```

If the child has one diagnosis, having an additional secondary diagnosis is associated with slightly higher social deficits (for most mABC scores) and a different pattern for the relationship between motor skills and social deficits

###Compare Average SRS Score for both of the Groups

####I think that one possible way to group this data would be to divide it into typically developing children, children with only Autism, Children with ADHD, and Children with both

####Single Diagnosis (either Autism or ADHD)
```{r}
p1 <- ggplot(data = singdia, aes(x = mABC_TotalStandardScore, y = SRS_TotalRawScore, color=PrimaryDiagnosis)) + geom_point() + geom_jitter() + geom_smooth(se=FALSE) +geom_smooth(data=singdia, aes(x = mABC_TotalStandardScore, y = SRS_TotalRawScore), size=1.25, color="green", se=FALSE)+theme(legend.position="bottom")
p1
```

####Multiple Diagnoses
```{r}
p2 <- ggplot(data = multdia, aes(x = mABC_TotalStandardScore, y = SRS_TotalRawScore, color=PrimaryDiagnosis)) + geom_point() + geom_jitter() + geom_smooth()+geom_smooth(data=multdia, aes(x = mABC_TotalStandardScore, y = SRS_TotalRawScore), size=1.25, color="green", se=FALSE)+theme(legend.position="bottom")
p2
```

####Typical Development
```{r}
p3 <- ggplot(data = nodia, aes(x = mABC_TotalStandardScore, y = SRS_TotalRawScore,colour = PrimaryDiagnosis)) + geom_point() + geom_jitter() + geom_smooth()+theme(legend.position="bottom")
p3
```

##Compare the Three Groups
```{r}
grid.arrange(p1, p2, p3, nrow=2)
```

####Single Diagnosis (either Autism or ADHD) mABC Components
```{r}
singdex<- ggplot(data = singdia, aes(x = mABC_ManualDexterity.Component.StandardScore, y = SRS_TotalRawScore, color=PrimaryDiagnosis)) + geom_point() + geom_jitter() + geom_smooth(se=FALSE) +xlab("Dexterity Subscore") +geom_smooth(data=singdia, aes(x = mABC_ManualDexterity.Component.StandardScore, y = SRS_TotalRawScore), size=1.25, color="green", se=FALSE)+theme(legend.position = "bottom")

singaimcat <- ggplot(data = singdia, aes(x = mABC_AimingAndCatching.Component.StandardScore, y = SRS_TotalRawScore, color=PrimaryDiagnosis, width=7)) + geom_point() + geom_jitter() + geom_smooth(se=FALSE)+ xlab("Aiming and Catching Subscore") +geom_smooth(data=singdia, aes(x = mABC_AimingAndCatching.Component.StandardScore, y = SRS_TotalRawScore), size=1.25, color="green", se=FALSE)+ theme(legend.position="none")

singbal <- ggplot(data = singdia, aes(x = mABC_Balance.Component.StandardScore, y = SRS_TotalRawScore, color=PrimaryDiagnosis, width=7)) + geom_point() + geom_jitter() + geom_smooth(se=FALSE)+ xlab("Balance Subscore") +geom_smooth(data=singdia, aes(x = mABC_Balance.Component.StandardScore, y = SRS_TotalRawScore), size=1.25, color="green", se=FALSE)+ theme(legend.position="none")

grid.arrange(singbal,singaimcat,singdex, nrow = 2)
```


####Multiple Diagnosis mABC Components

```{r}
multdex <-ggplot(data = multdia, aes(x = mABC_ManualDexterity.Component.StandardScore, y = SRS_TotalRawScore, color=PrimaryDiagnosis)) + geom_point() + geom_jitter() + geom_smooth()+theme(legend.position="bottom")+xlab("Dexterity Subscore")+geom_smooth(data=multdia, aes(x = mABC_ManualDexterity.Component.StandardScore, y = SRS_TotalRawScore), size=1.25, color="green", se=FALSE)
  
multaimcat <-ggplot(data = multdia, aes(x = mABC_AimingAndCatching.Component.StandardScore, y = SRS_TotalRawScore, color=PrimaryDiagnosis)) + geom_point() + geom_jitter() + geom_smooth()+theme(legend.position="none")+ xlab("Aiming and Catching Subscore")+geom_smooth(data=multdia, aes(x =  mABC_AimingAndCatching.Component.StandardScore, y = SRS_TotalRawScore), size=1.25, color="green", se=FALSE)
  
  
multbal <- ggplot(data = multdia, aes(x = mABC_Balance.Component.StandardScore, y = SRS_TotalRawScore, color=PrimaryDiagnosis)) + geom_point() + geom_jitter() + geom_smooth()+theme(legend.position="none") + xlab("Balance Subscore")+geom_smooth(data=multdia, aes(x =  mABC_Balance.Component.StandardScore, y = SRS_TotalRawScore), size=1.25, color="green", se=FALSE)
  

grid.arrange(multbal,multaimcat,multdex, nrow = 2)
```

####Typical Development mABC Components

```{r}
typdex<-ggplot(data = nodia, aes(x = mABC_ManualDexterity.Component.StandardScore, y = SRS_TotalRawScore)) + geom_point() + geom_jitter() + geom_smooth()+theme(legend.position="none")+xlab("Dexterity Subscore")

typaimcat<-ggplot(data = nodia, aes(x = mABC_ManualDexterity.Component.StandardScore, y = SRS_TotalRawScore)) + geom_point() + geom_jitter() + geom_smooth()+theme(legend.position="none")+xlab("Dexterity Subscore")

typbal<-ggplot(data = nodia, aes(x = mABC_ManualDexterity.Component.StandardScore, y = SRS_TotalRawScore)) + geom_point() + geom_jitter() + geom_smooth()+theme(legend.position="none")+xlab("Dexterity Subscore")

grid.arrange(typdex, typaimcat, typbal, nrow=2)
```