---
title: "Project Proposal & Preliminary Findings: "
subtitle: "KKI Study on Motor Skills and Social Deficits in Children"
author: "Allison McCarty,  Nick Janes, Thomas Mobely"
date: "11/13/2019"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

```{r, include=FALSE}
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(gridExtra)
library(ggthemes)
library(reportr)

kki.demo <- read.csv("KKI_demographicInfo.csv")
kki.social <- read.csv("KKI_SocialResponsivenessScaleQuestionnaire.csv")
kki.movement <- read.csv("KKI_movementAssessmentBatteryforChildren.csv")
#Fix odd ID col
colnames(kki.demo)[1] <- "ID"
colnames(kki.social)[1] <- "ID"
colnames(kki.movement)[1] <- "ID"
kki <- merge(kki.demo,kki.social)
kki <- merge(kki,kki.movement)
#Now have all the data in a huge data frame
#Removing N/A values. Mostly removing second visits.
kki <- na.omit(kki)
kki2 <- filter(kki,PrimaryDiagnosis != "None",SecondaryDiagnosis ==c("Yes", "No"))
multdia <- filter(kki, PrimaryDiagnosis != "None", SecondaryDiagnosis =="Yes")
singdia <- filter(kki, PrimaryDiagnosis != "None" , SecondaryDiagnosis == "No")
nodia <- filter(kki, PrimaryDiagnosis =="None" , SecondaryDiagnosis =="No")
```
## Introduction: Study Details & Relevant Background Information

- Autism (ASD) and Attention Deficit/Hyperactivity Disorder (ADHD) are frequently observed in children, but are still not fully understood. 

- Researchers at Kennedy Krieger Institute(KKI)  ran a study on children with ASD, ADHD, and typically developing children to study differences in social deficits, motor impairment, intelligence.
    - It is known that children with developmental disorders suffer from motor impairment concurrent with social deficits       associated with their respective diagnoses(1,2).  The extent and cause of this correlation merits further research.

- Metrics and Scales Used to Quantify Symptom Severity in Children:
    1. Social Responsiveness Scale Questionnaire (SRS): measures severity of social deficits in children (higher scores indicate more social deficits)
    2. Movement Assessment Battery for Children (mABC): measures motor impairments in children.  Three subcategories for assessment include dexterity, aiming and catching, and balance summed to obtain the mABC Total Standard Score
    3. Wechsler Intelligence Scale for Children (WISC): measures intelligence in children.  Two versions of the test quantify vocabulary, information, picture concepts, etc.

## Introduction: Areas of Particular Interest
###Data and Analysis

- We want to investigate the overall association between social deficits and motor skills for each diagnosis group.


    <large>> Primary Endpoint: What motor skills are associated with SRS scores for each diagnosis group.</large>
  
    <large>> Secondary Endpoint: What factors influence this association?</large>
    
- Data Collected for Investigation and Analysis:
    - Age
    - Handedness (quantified by Edinburgh Handedness Integer)
    - Gender
    - Primary and Secondary Diagnosis
    - SRS Total and Subscores
    - mABC Total and Subscores
    - WISC Score
    

##Relationship Between SRS Total Score and Motor Assessment Score
```{r}
ggplot(data = kki, aes(x = mABC_TotalStandardScore, y = SRS_TotalRawScore,colour = PrimaryDiagnosis)) + geom_point(alpha = .3) + geom_jitter() + geom_smooth()+ylab("SRS Total Score")+xlab("Motor Assessment Score")+ggtitle("Overall Association Between Social Deficits and Motor Ability")+ theme(legend.position = 'bottom')

```

##Relationship Between SRS Total Score and Motor Assessment Sub-Scores

```{r, include=FALSE}
SRS.Manual <- ggplot(data = kki, aes(x = mABC_ManualDexterity.Component.StandardScore, y = SRS_TotalRawScore, colour = PrimaryDiagnosis)) + geom_point(alpha = .3) + geom_jitter() + geom_smooth()+ theme(legend.position = 'none')+ylab("SRS Total Score")+xlab("Manual Dexterity Score")+ggtitle("Association between Manual Dexterity and Socail Deficits")
```

```{r, include=FALSE}
SRS.AimAndCatch <- ggplot(data = kki, aes(x = mABC_AimingAndCatching.Component.StandardScore, y = SRS_TotalRawScore, colour = PrimaryDiagnosis)) + geom_point(alpha = .3) + geom_jitter() + geom_smooth()+ theme(legend.position = 'none')+ylab("SRS Total Score")+xlab("Aiming and Catching Score")+ggtitle("Association between Aiming and Catching and Socail Deficits")
```

```{r, include=FALSE}
#Really interesting! 
#The better the ADHD kids were at balance the better their social deficit was! Let's investigate the relationship between age and balance
SRS.Balance <- ggplot(data = kki, aes(x = mABC_Balance.Component.StandardScore, y = SRS_TotalRawScore, colour = PrimaryDiagnosis)) + geom_point(alpha = .3) + geom_jitter() + geom_smooth() + theme(legend.position = 'bottom')+ylab("SRS Total Score")+xlab("Balance Score")+ggtitle("Association Between Balance and Social Deficits")
```

```{r}
grid.arrange(SRS.Manual,SRS.AimAndCatch,SRS.Balance, nrow = 2)
```

##Does Intelligence(measured by WISC) influence Social Deficits or Motor Skills?
```{r, include=FALSE}
wisc <- read.csv("KKI_WechslerIntelligenceScaleforChildren.csv")
#wisc <- na.omit(wisc)
wisc4 <- filter(wisc, WISC_VERSION == 4)
wisc5 <- filter(wisc, WISC_VERSION == 5)
wisc4GAI <- read.csv("wisc4GAI_lookup.csv")
wisc5GAI <- read.csv("wisc5GAI_lookup.csv")
kki.demo <- read.csv("KKI_demographicInfo.csv")
kki.social <- read.csv("KKI_SocialResponsivenessScaleQuestionnaire.csv")
kki.movement <- read.csv("KKI_movementAssessmentBatteryforChildren.csv")
wisc4$score <- rowSums(wisc4[,5:10])
wisc4$WISC5_Vocabulary <- NULL
wisc4$WISC5_Similarities <- NULL
wisc4$WISC5_BlockDesign <- NULL
wisc4$WISC5_MatrixReasoning <- NULL
wisc4$WISC5_FigureWeights <- NULL
wisc4 <- na.omit(wisc4)
wisc5$score <- rowSums(wisc5[,11:15])
wisc5$WISC4_Vocabulary <- NULL
wisc5$WISC4_Similarities <- NULL
wisc5$WISC4_BlockDesign <- NULL
wisc5$WISC4_MatrixReasoning <- NULL
wisc5$WISC4_FigureWeights <- NULL
wisc5$WISC4_Information <- NULL
wisc5$WISC4_PictureConcepts <- NULL
wisc5 <- na.omit(wisc5)
```


```{r, include=FALSE}
#Fix odd ID col
colnames(kki.demo)[1] <- "ID"
colnames(kki.social)[1] <- "ID"
colnames(kki.movement)[1] <- "ID"
kki <- merge(kki.demo,kki.social)
kki <- merge(kki,kki.movement)
#Now have all the data in a huge data frame
#Removing N/A values. Mostly removing second visits.
kki <- na.omit(kki)
#Combine kki with WISC version 4 data
kki.wisc4 <- full_join(kki, wisc4, by = 'ID')
kki.wisc4 <- filter(kki.wisc4, WISC_VERSION == 4)
#Combine kki with WISC version 5 data
kki.wisc5 <- full_join(kki, wisc5, by = 'ID')
kki.wisc5 <- filter(kki.wisc5, WISC_VERSION == 5)
kki2 <- filter(kki,PrimaryDiagnosis != "None",SecondaryDiagnosis ==c("Yes", "No"))
multdia <- filter(kki, PrimaryDiagnosis != "None", SecondaryDiagnosis =="Yes")
singdia <- filter(kki, PrimaryDiagnosis != "None" , SecondaryDiagnosis == "No")
nodia <- filter(kki, PrimaryDiagnosis =="None" , SecondaryDiagnosis =="No")
```

```{r, include=FALSE}
intel1 <-ggplot(data = kki.wisc4, aes(x = score, y = SRS_TotalRawScore, colour = PrimaryDiagnosis)) + geom_point() + ggtitle("WISC Version 4: SRS Score vs. WISC Score") + labs(x = "WISC Score", y = "SRS Total Score") + geom_smooth()+theme(legend.position = 'none')

intel2 <- ggplot(data = kki.wisc5, aes(x = score, y = SRS_TotalRawScore, colour = PrimaryDiagnosis)) + geom_point() + ggtitle("WISC Version 5: SRS Score vs. WISC Score") + labs(x = "WISC Score", y = "SRS Total Score") + geom_smooth()+theme(legend.position = 'none')

intel3 <-ggplot(data = kki.wisc4, aes(x = score, y = mABC_TotalStandardScore, colour = PrimaryDiagnosis)) + geom_point() + ggtitle("WISC Version 4: mABC Score vs. WISC Score") + labs(x = "WISC Score", y = "mABC Total Score") + geom_smooth()+theme(legend.position = 'none')

intel4 <-ggplot(data = kki.wisc5, aes(x = score, y = mABC_TotalStandardScore, colour = PrimaryDiagnosis)) + geom_point() + ggtitle("WISC Version 5: mABC Score vs. WISC Score") + labs(x = "WISC Score", y = "mABC Score") + geom_smooth()+theme(legend.position = 'bottom')
```

```{r}
grid.arrange(intel1, intel2, intel3, intel4, nrow=2)
```

## Does having multiple diagnoses influence Social Deficits or Motor Skills?

```{r, include=FALSE}
p1 <- ggplot(data = singdia, aes(x = mABC_TotalStandardScore, y = SRS_TotalRawScore, color=PrimaryDiagnosis)) + geom_point() + geom_jitter() + geom_smooth(se=FALSE) +geom_smooth(data=singdia, aes(x = mABC_TotalStandardScore, y = SRS_TotalRawScore), size=1.25, color="green", se=FALSE)+theme(legend.position="bottom") +
  ggtitle("Single Diagnosis Group 
(either ASD or ADHD)") +ylab("SRS Total Score")+xlab("Motor Assessment Score")
p1
```


```{r, include=FALSE}
p2 <- ggplot(data = multdia, aes(x = mABC_TotalStandardScore, y = SRS_TotalRawScore, color=SecondaryDiagnosis)) + geom_point() + geom_jitter() + geom_smooth(color="green")+theme(legend.position="none")+
  ggtitle("Multiple Diagnosis Group
(both ADHD and Autism)")+ylab("SRS Total Score")+xlab("Motor Assessment Score")
p2
```


```{r, include=FALSE}
p3 <- ggplot(data = nodia, aes(x = mABC_TotalStandardScore, y = SRS_TotalRawScore,colour = PrimaryDiagnosis)) + geom_point() + geom_jitter() + geom_smooth(color="green")+theme(legend.position="bottom")+ggtitle("Typical Development Group")+ylab("SRS Total Score")+xlab("Motor Assessment Score")+theme(legend.position="none")
p3
```


```{r}
grid.arrange(p1, p2, p3, nrow=2)
```

##Does Handedness influence Social Deficits or Motor Skills?

```{r, include=FALSE}
kki.hand <-  read.csv("KKI_handedness.csv")
kki.demo <- read.csv("KKI_demographicInfo.csv")
kki.move <- read.csv("KKI_movementAssessmentBatteryforChildren.csv")
kki.srs <- read.csv("KKI_SocialResponsivenessScaleQuestionnaire.csv")
colnames(kki.hand)[1] <- "ID"
colnames(kki.demo)[1] <- "ID"
colnames(kki.move)[1] <- "ID"
colnames(kki.srs)[1] <- "ID"
kki.sub1 <- merge(kki.hand,kki.demo)
kki.sub2 <- merge(kki.move,kki.srs)
kki <- merge(kki.sub1,kki.sub2)

kki.need <- cbind(kki[1],kki[3],kki[12],kki[5],kki[16],kki[19],kki[23],kki[27])

left <- subset(kki.need,kki$EdinburghHandedness_Integer < 0)
left <- subset(left,left$PrimaryDiagnosis == "Autism")
right <- subset(kki.need,kki$EdinburghHandedness_Integer > 0)
right <- subset(right,right$PrimaryDiagnosis == "Autism")

kki$RightHanded <- kki$EdinburghHandedness_Integer > 0
kki.need$RightHanded <- kki$EdinburghHandedness_Integer >= 0


hand1 <-ggplot(data = na.omit(kki.need),aes(x = RightHanded, y = mABC_AimingAndCatching.Component.StandardScore, colour = PrimaryDiagnosis)) + geom_boxplot() + facet_wrap(~PrimaryDiagnosis)+theme(legend.position="none")+ylab("Aiming and Catching Score")

hand2 <- ggplot(data = na.omit(kki.need),aes(x = RightHanded, y = mABC_Balance.Component.StandardScore, colour = PrimaryDiagnosis)) + geom_boxplot() + facet_wrap(~PrimaryDiagnosis)+theme(legend.position="none")+ylab("Balance Score")

hand3 <- ggplot(data = na.omit(kki.need),aes(x = RightHanded, y = mABC_ManualDexterity.Component.StandardScore, colour = PrimaryDiagnosis)) + geom_boxplot() + facet_wrap(~PrimaryDiagnosis)+theme(legend.position="bottom")+ylab("Manual Dexterity Score")

```


```{r}
grid.arrange(hand1, hand2, hand3, nrow=2)
```

##Does Age influence Social Deficits or Motor Skills?


```{r, include=FALSE}
Move.Age <- ggplot(data = kki, aes(x = mABC_AGE, y = mABC_TotalStandardScore, colour = PrimaryDiagnosis)) + geom_point(alpha = .3) + geom_jitter() + geom_smooth() + theme(legend.position = 'none')+xlab("Age(years)")+ylab("mABC Total Score")
#low correlation between age and motor skills in all 
# groups
```

```{r, include=FALSE}
SRS.Age <- ggplot(data = kki, aes(x = SRS_AGE, y = SRS_TotalRawScore, colour = PrimaryDiagnosis)) + geom_point(alpha = .3) + geom_jitter() + geom_smooth() + theme(legend.position = 'none')+xlab("Age(years)")+ylab("SRS Total Score")
#again no real relationship
#NEXT CHUNK:I don't know a good way of putting the gird together
# and the legend so somebody might wanna try that
```

```{r}
grid.arrange(Move.Age,SRS.Age, nrow = 1)

```



##Summary of Preliminary and Research Proposal

-

##Citations

1. Green,  D.,  Charman,  T.,  Pickles,  A.,  Chandler,  S.,  Loucas,  T.,Simonoff,  E.,  et  al.  (2009).  Impairment  in  movement  skills  ofchildren  with  autistic  spectrum  disorders. Developmental  Medicine and Child Neurology, 51(4), 311–316.

2. Ament, K., Mejia, A., E., et al. (2014). Evidence for specificity of motor impairments in catching and balance in children with autism. Journal of autism and developmental disorders, 45(3), pp.742-751.




