---
title: "Linear Models S431 Final Presentation:"
subtitle: "Exploring the Relationship between Motor Skills and Social Deficits in Children"
author: "Alli, Nick, Thomas"
date: "12/4/2019"
output: 
  slidy_presentation
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r,echo = FALSE,include=FALSE}
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(gridExtra)
library(ggthemes)
library(reportr)
library(knitr)
library(splines)
library(ggpubr)
```

```{r, echo=FALSE,include=FALSE}
kki.demo <- read.csv("Data/KKI_demographicInfo.csv")
kki.social <- read.csv("Data/KKI_SocialResponsivenessScaleQuestionnaire.csv")
kki.movement <- read.csv("Data/KKI_movementAssessmentBatteryforChildren.csv")
kki.hand <- read.csv("Data/KKI_handedness.csv")
wisc <- read.csv("Data/KKI_WechslerIntelligenceScaleforChildren.csv")
wisc4GAI <- read.csv("Data/wisc4GAI_lookup.csv")
wisc5GAI <- read.csv("Data/wisc5GAI_lookup.csv")
colnames(kki.demo)[1] <- "ID"
colnames(kki.social)[1] <- "ID"
colnames(kki.movement)[1] <- "ID"
colnames(kki.hand)[1] <- "ID"
colnames(wisc)[1] <- "ID"
colnames(wisc4GAI)[1] <- "ID"
colnames(wisc5GAI)[1] <- "ID"
wisc4 <- filter(wisc, WISC_VERSION == 4)
wisc5 <- filter(wisc, WISC_VERSION == 5)
wisc4 <- wisc4[-c(11:15)]
wisc5 <- wisc5[-c(5:10)]
wisc4 <- filter(wisc4, visit == 1)
wisc5 <- filter(wisc5, visit == 1)
#Calculate composite scores for both WISC versions
wisc4$score <- rowSums(wisc4[,5:10])
wisc5$score <- rowSums(wisc5[,5:9])
#Filter out NA scores
wisc4 <- na.omit(wisc4, cols = "score")
wisc5 <- na.omit(wisc5, cols = "score")
#Convert composite scores to GAI scores for WISC version 4
composite4 = rep(0,length(wisc4[,1]))
for (i in 1:length(wisc4[,1])) {
  composite4[i] <- wisc4GAI[wisc4[i,11]-5,2]
}
wisc4$GAI <- composite4
#Convert composite scores to GAI scores for WISC version 5
composite5 = rep(0,length(wisc5[,1]))
for (i in 1:length(wisc5[,1])) {
  composite5[i] <- wisc5GAI[wisc5[i,10]-4,2]
}
wisc5$GAI <- composite5
#Remove composite scores
wisc4 <- wisc4[-c(2:11)]
wisc5 <- wisc5[-c(2:10)]
#Combine WISC data
wisc.combo <- full_join(wisc4, wisc5, by = c('ID', 'GAI'))
```

```{r,echo=FALSE,include=FALSE}
kki <- merge(kki.demo,kki.social)
kki <- merge(kki,kki.movement)
kki <- merge(kki,wisc.combo)
kki <- merge(kki,kki.hand)
kki.all <- cbind(kki,rep(1,nrow(kki)))
kki$Hand <- cut(kki$EdinburghHandedness_Integer,c(-1.01,-.5,.5,1.01),labels = c("Left","Mixed","Right"))
## Making standard Scores
kki.V1 <- subset(kki,kki$SRS_VERSION == 1)
kki.V2 <- subset(kki,kki$SRS_VERSION == 2)

Two_to_One <- function(from,target) {
  y.bar = mean(target);y.sd = sd(target);
  x.bar = mean(from);x.sd = sd(from);
  right.half = (y.bar - (y.sd/x.sd) * x.bar)
  left.half = y.sd/x.sd
  return (left.half * from + right.half)
}
kki.V1$Standard <- kki.V1$SRS_TotalRawScore
kki.V2$Standard <- Two_to_One(kki.V2$SRS_TotalRawScore,kki.V1$SRS_TotalRawScore)
kki <- rbind(kki.V1,kki.V2)
kki$sqrt.Standard <- sqrt(kki$Standard)
## 555
#Now have all the data in a huge data frame
#Removing N/A values. Mostly removing second visits.
kki <- subset(kki,!is.na(kki$SRS_TotalRawScore))
kki <- subset(kki, !is.na(kki$mABC_TotalStandardScore))
kki <- subset(kki,kki$mABC_AGE<14)
kki$Hand <- cut(kki$EdinburghHandedness_Integer,c(-1.01,-.5,.5,1.01),labels = c("Left","Mixed","Right"))
kki2 <- filter(kki,PrimaryDiagnosis != "None",SecondaryDiagnosis ==c("Yes", "No"))
multdia <- filter(kki, PrimaryDiagnosis != "None", SecondaryDiagnosis =="Yes")
singdia <- filter(kki, PrimaryDiagnosis != "None" , SecondaryDiagnosis == "No")
nodia <- filter(kki, PrimaryDiagnosis =="None" , SecondaryDiagnosis =="No")
kki.adhd <- subset(kki,PrimaryDiagnosis == "ADHD")
kki.adhd <- subset(kki.adhd,ADHD_Subtype != "Hyperactive/Impulsive")
kki.adhd1 <- subset(kki.adhd,kki.adhd$SRS_VERSION == 1)
kki.adhd2 <- subset(kki.adhd, kki.adhd$SRS_VERSION == 2)
```


## Study Details & Relevant Background Information

- Autism (ASD) and Attention Deficit/Hyperactivity Disorder (ADHD) are frequently observed in children, but are still not fully understood. 


    - It is known that children with developmental disorders suffer from motor impairment concurrent with social deficits       associated with their respective diagnoses(1,2).  The extent and cause of this correlation merits further research.

- Kennedy Krieger Institute(KKI) Study: Data Collected for Investigation and Analysis: 

         Social Responsiveness Scale Questionnaire (SRS), Movement Assessment Battery 
         for Children (mABC), Wechsler Intelligence Scale for Children (WISC), Gender, 
         Age,Handedness (quantified by Edinburgh Handedness Integer), Primary and 
         Secondary Diagnosis

- We want to investigate the overall association between social deficits and motor skills for each diagnosis group.

    - Primary Endpoint: What motor skills are associated with SRS scores for each diagnosis group.
  
    - Secondary Endpoint: What factors influence this association?


## Methods: Cleaning Data

```{r,include = FALSE,echo = FALSE}
kki.asd <- subset(kki,kki$PrimaryDiagnosis == 'Autism')
kki.adhd <- subset(kki,kki$PrimaryDiagnosis == 'ADHD')
kki.td <- subset(kki,kki$PrimaryDiagnosis == 'None')
diagnosis <- c("ASD","ADHD","TD")
counts <- c(nrow(kki.asd),nrow(kki.adhd),nrow(kki.td))
age <- c(mean(kki.asd$mABC_AGE),mean(kki.adhd$mABC_AGE),mean(kki.td$mABC_AGE))
kki.asd.male <- subset(kki.asd,kki.asd$Gender == 'M')
kki.asd.female <- subset(kki.asd,kki.asd$Gender == 'F')
kki.adhd.male <- subset(kki.adhd,kki.adhd$Gender == 'M')
kki.adhd.female <- subset(kki.adhd,kki.adhd$Gender == 'F')
kki.td.male <- subset(kki.td, kki.td$Gender == 'M')
kki.td.female <- subset(kki.td,kki.td$Gender == 'F')
maleCount <- c(nrow(kki.asd.male),nrow(kki.adhd.male),nrow(kki.td.male))
femaleCount <- c(nrow(kki.asd.female),nrow(kki.adhd.female),nrow(kki.td.female))
rightCount <- c(length(which(kki.asd$Hand == "Right")),length(which(kki.adhd$Hand == "Right")),length(which(kki.td$Hand == "Right")))
leftCount <- c(length(which((kki.asd$Hand == "Left"))),length(which(kki.adhd$Hand == "Left")),length(which(kki.td$Hand == "Left")))
mixedCount <- c(length(which((kki.asd$Hand == "Mixed"))),length(which((kki.adhd$Hand == "Mixed"))),length(which((kki.td$Hand == "Mixed"))))
table1 <- cbind(diagnosis,counts,round(age,4),maleCount,femaleCount,leftCount,mixedCount,rightCount)
FinalTable <- kable(table1,col.names = c("Primary","Count","Avg Age","Males","Females","Left","Mixed","Right"))
agehist <-ggplot(data = kki.all,aes(x = mABC_AGE, colour = PrimaryDiagnosis, fill=PrimaryDiagnosis)) + geom_histogram(fill="white")+ylab("Age (Years)")+theme_classic() + geom_vline(xintercept = 14.5)
```

- Removed Subjects that were under 14, since 94% subjects diagnosed with Autism or ADHD were under 14. Reduced total observations from 832 to 755

```{r, fig.height=2, fig.width=5, fig.align="center", include=TRUE}
agehist
```

- Then Removed Observations that had N/A values for either SRS_TotalRawScore or mABC_TotalStandardScore which furthur reduced the total observations to 568 Observations



- Converted all the WISC scores to the GAI scale for comparability across versions

```{r, height=1}
FinalTable
```




## Methods: Exploring Motor Skill's Association with Social Deficits


```{r,echo = FALSE}
graph.dat <- subset(kki,kki$mABC_TotalStandardScore <12)
total <- ggplot(data = graph.dat, aes(x = mABC_TotalStandardScore, y = Standard,colour = PrimaryDiagnosis)) + geom_point(alpha = .3) + geom_jitter() + geom_smooth()+ylab("SRS Total Score")+xlab("Motor Assessment Score")+ggtitle("Overall Motor Ability and Social Deficits")+ theme(legend.position = 'none')
```

```{r,echo = FALSE,include=FALSE}
SRS.Manual <- ggplot(data = kki, aes(x = mABC_ManualDexterity.Component.StandardScore, y = Standard, colour = PrimaryDiagnosis)) + geom_point(alpha = .3) + geom_jitter() + geom_smooth()+ theme(legend.position = 'none')+ylab("SRS Total Score")+xlab("Manual Dexterity Score")+ggtitle("Manual Dexterity and Social Deficits")
```

```{r,echo = FALSE,include=FALSE}
SRS.AimAndCatch <- ggplot(data = kki, aes(x = mABC_AimingAndCatching.Component.StandardScore, y = Standard, colour = PrimaryDiagnosis)) + geom_point(alpha = .3) + geom_jitter() + geom_smooth()+ theme(legend.position = 'none')+ylab("SRS Total Score")+xlab("Aiming and Catching Score")+ggtitle("Aiming & Catching and Social Deficits")
```

```{r,echo = FALSE,include=FALSE}
#Really interesting! 
#The better the ADHD kids were at balance the better their social deficit was! Let's investigate the relationship between age and balance
graph.dat <- subset(kki, kki$mABC_Balance.Component.StandardScore < 15)
SRS.Balance <- ggplot(data = graph.dat, aes(x = mABC_Balance.Component.StandardScore, y = Standard, colour = PrimaryDiagnosis)) + geom_point(alpha = .3) + geom_jitter() + geom_smooth() + theme(legend.position = 'none')+ylab("SRS Total Score")+xlab("Balance Score")+ggtitle("Balance and Social Deficits")
```

```{r,echo = FALSE, fig.height=5, fig.align="center", fig.width=8}
ggarrange(SRS.Manual, SRS.AimAndCatch, SRS.Balance, total, common.legend = TRUE, legend="bottom")
```

  - Key Takeaway: The association between motor skills and social deficits is different for each of the subcomponents of the mABC, and differ for children with different primary diagnoses.



 
## Methods: Other Variables

```{r, include=FALSE}
gaisrs1 <-ggplot(data = kki, aes(x = GAI, y = Standard,colour = PrimaryDiagnosis)) + geom_point() + geom_smooth() + ylab("SRS Total Score")+xlab("GAI Test Score")+ggtitle("GAI Score and Social Deficits")+ theme(legend.position="none")
gaisrs2 <-ggplot(data = kki, aes(x = GAI, y = mABC_TotalStandardScore,colour = PrimaryDiagnosis)) + geom_point() + geom_jitter() + geom_smooth() + ylab("Total Motor Score")+xlab("GAI Test Score")+ggtitle("GAI Score and Motor Skills")+ theme(legend.position="none")
```


```{r, fig.height=4, fig.width=5, include=FALSE}
ggarrange(gaisrs1, gaisrs2, nrow=1, common.legend = TRUE, legend="bottom")
```

 

```{r, include=FALSE}
kki$Hand <- cut(kki$EdinburghHandedness_Integer,c(-1.01,-.5,.5,1.01),labels = c("Left","Mixed","Right"))
hand1 <- ggplot(data = na.omit(kki),aes(group = Hand,y = mABC_TotalStandardScore, colour = Hand)) + geom_boxplot() + facet_wrap(na.omit(kki)$PrimaryDiagnosis)+ theme(legend.position="none")+ ylab("Total Motor Score")+ggtitle("Handedness & Motor Skills")
hand2 <-ggplot(data = na.omit(kki),aes(group = Hand,y = Standard, colour = Hand)) + geom_boxplot() + facet_wrap(na.omit(kki)$PrimaryDiagnosis)+ theme(legend.position="none")+ ylab("SRS Total Score")+ggtitle("Handedness & Social Deficits")
```

```{r, fig.height=4, fig.width=11, fig.align="left"}
ggarrange(gaisrs1, gaisrs2, hand2, hand1, nrow=1,common.legend = TRUE, legend="bottom")
```

- Key Takeaways: 

  > There seems to be a strong negative correlation between intelligence and motor skills, but moderate positive correlation between intelligence and social deficit.
  
  > The different groups seem to have similiar Social Deficits, but vastly different Motor SKills.

## Proposed Models

$$
\text{E[Social Deficit Score]} = \textbf{X} * \boldsymbol{\beta} \\
where: \\
\boldsymbol{\beta} = \begin{bmatrix}
            \text{Intercept} \\
            \text{Balance Score} \\
            \text{Aiming And Catching Score} \\
            \text{EdinburgHandedness_Integer} \\
            \text{GAI} \\
            \text{Gender Male} \\
            \text{mABC_AGE} 
            \end{bmatrix}
$$

```{r,include=FALSE}
kki.1 <- subset(kki,kki$SRS_VERSION == 1)
kki.2 <- subset(kki,kki$SRS_VERSION == 2)
t.test(kki.1$SRS_TotalRawScore,kki.2$SRS_TotalRawScore)
```

Seperate SRS Versions from Welch t-test:

$$
H_0: \text{E[SRS Scores Version 1] = E[SRS Scores Version 2]} \\
H_1: \text{E[SRS Scores Version 1]} \space \neq \text{E[SRS Scores Version 2]} \\
\text{Under the null we got a t-statistc of -6.03 from a t-distribution of 417.38 } \\
\text{that has a p-value of} \space 3.511 \times 10^{-9} \\ \space \text{Under} \space \alpha = .05 \space \text{we reject the null}
\text{ and believe that there is a true difference in the mean}
$$

## Resutls: Typically Developing

```{r}
nodia1 <- subset(nodia,nodia$SRS_VERSION == 1)
nodia2 <- subset(nodia,nodia$SRS_VERSION == 2)
lm.nodia1 <- lm(Standard ~ mABC_AimingAndCatching.Component.StandardScore + mABC_Balance.Component.StandardScore + EdinburghHandedness_Integer + GAI  + Gender + mABC_AGE, data = nodia1)
lm.nodia2 <- lm(Standard ~ mABC_AimingAndCatching.Component.StandardScore + mABC_Balance.Component.StandardScore + EdinburghHandedness_Integer + GAI  + Gender + mABC_AGE, data = nodia2)
```


$$
\text{E[Social Deficit Score]} = \textbf{X} * \boldsymbol{\beta} \\
where: \\
\boldsymbol{\beta} = \begin{bmatrix}
            \text{Intercept} \\
            \text{Balance Score} \\
            \text{Aiming And Catching Score} \\
            \text{EdinburgHandedness_Integer} \\
            \text{GAI} \\
            \text{Gender Male} \\
            \text{mABC_AGE} 
            \end{bmatrix}
=
\begin{bmatrix}
43.52523 \\
-0.14406 \\
0.01693  \\
-1.46571  \\
0.05956   \\
-1.12175  \\
-0.09471
\end{bmatrix}
$$


$$
P-Value=0.3062
$$
$$
Adjusted R-Squared:  0.01153
$$





## Results: ADHD

```{r,echo=FALSE,include=FALSE}
lm.adhd <- lm(sqrt.Standard ~ mABC_AimingAndCatching.Component.StandardScore + mABC_Balance.Component.StandardScore + EdinburghHandedness_Integer + GAI + ADHD_Subtype + Gender + mABC_AGE, data = kki.adhd)
lm.basic <- lm(sqrt.Standard ~ mABC_AimingAndCatching.Component.StandardScore + mABC_Balance.Component.StandardScore + mABC_ManualDexterity.Component.StandardScore, data = kki.adhd)
```

$$
\boldsymbol{\beta} = \begin{bmatrix}
            \text{Intercept} \\
            \text{Balance Score} \\
            \text{Aiming And Catching Score} \\
            \text{EdinburgHandedness_Integer} \\
            \text{GAI} \\
            \text{ADHD Subtype Inattentive} \\
            \text{Gender Male} \\
            \text{mABC_AGE} 
\end{bmatrix}
=
\begin{bmatrix}
7.2115 \\
-0.0408 \\
-0.0646  \\
0.3292  \\
0.0085   \\
-1.266 \\
-0.175 \\
-0.049 
\end{bmatrix}
=
\begin{bmatrix}
6.11 - 8.24 \\
-0.08 - (-0.0042)  \\
-0.1033 - (-0.0433) \\
0.1356 - 0.5182  \\
0.00261 - 0.0144   \\
-1.563711 - (-0.9777) \\
-0.414 - (0.078) \\
-0.1403 - 0.0456
\end{bmatrix}
$$

```{r,include= FALSE}
studentize <- function(model) {
  n = length(fitted(model))
  p = length(model$coefficients)
  s = (1/(n-p-1)) * sum(residuals(model)^2)
  residuals(model)/ (s * (sqrt(1-influence(model)$hat)))
}
student.lmadhd <- studentize(lm.adhd)
leverage.lmadhd <- influence(lm.adhd)$hat
```


```{r, fig.height=3, fig.width=4}
ggplot(data = NULL,aes(x = fitted(lm.adhd), y = residuals(lm.adhd))) + geom_point() + geom_hline( yintercept = 0) + geom_smooth() +
  ggtitle("Residual Plot for ADHD Model(1)") + xlab("Fitted Values") + ylab("Residual Values")+theme_classic()
ggplot(data = NULL,aes(x = leverage.lmadhd, y = student.lmadhd)) + geom_point() + geom_hline(y = 0)
qqnorm(lm.adhd$residuals)
```

```{r}
CI_adhd <- confint(lm.adhd, c('mABC_AimingAndCatching.Component.StandardScore', 'mABC_Balance.Component.StandardScore', 'EdinburghHandedness_Integer', 'GAI', 'ADHD_SubtypeInattentive', 'GenderM', 'mABC_AGE'))
CI_LB_adhd <- CI_adhd[1]
CI_UB_adhd <- CI_adhd[2]
```


```{r,include = FALSE,echo = FALSE}
anova(lm.basic,lm.adhd)
```

$$
H_0: \text{Just Balance and Catching is Sufficient} \\
H_1: \text{Full Model is Sufficient} \\
\text{Under the null we got an F-statistic of 5.905 where} \ F \sim F_{4,165} \\
\text{which gives us a p-value of .00018 with an} \ \alpha = .05 \ \text{we reject the null} \\
\text{and believe that we should keep every coefficient in the full model}
$$



## Results: ASD

```{r}
aut <- subset(kki, kki$PrimaryDiagnosis == "Autism")
lm.aut <- lm(Standard ~ mABC_AimingAndCatching.Component.StandardScore + mABC_Balance.Component.StandardScore + EdinburghHandedness_Integer + GAI + ADHD_Subtype + Gender + mABC_AGE, data = aut)
summary(lm.aut)
CI_aut <- confint(lm.aut, c('mABC_AimingAndCatching.Component.StandardScore', 'mABC_Balance.Component.StandardScore', 'EdinburghHandedness_Integer', 'GAI', 'ADHD_SubtypeInattentive', 'GenderM', 'mABC_AGE'))
CI_LB_aut <- CI_aut[1]
CI_UB_aut <- CI_aut[2]
```

$$
\boldsymbol{\beta} = \begin{bmatrix}
            \text{Intercept} \\
            \text{Balance Score} \\
            \text{Aiming And Catching Score} \\
            \text{EdinburgHandedness_Integer} \\
            \text{GAI} \\
            \text{ADHD Subtype Inattentive} \\
            \text{Gender Male} \\
            \text{mABC_AGE} 
\end{bmatrix}
=
\begin{bmatrix}
54.25094 \\
0.37661 \\
-0.02558  \\
4.40532  \\
0.14978   \\
11.15424 \\
-6.57359 \\
 1.22425 
\end{bmatrix}
=
\begin{bmatrix}
18.95858 - 89.5433 \\
-1.52811116 - 2.2813354  \\
-1.84658543 - 1.7954352 \\
-4.30532586 - 13.1159718  \\
-0.05164134 - 0.3512058   \\
-40.23481750 - 62.5433065 \\
-19.03641089 - 5.8892330 \\
-2.07171924 - 4.5202255
\end{bmatrix}
$$


```{r, fig.height=3, fig.width=4}
ggplot(data = NULL,aes(x = fitted(lm.aut), y = residuals(lm.aut))) + geom_point() + geom_hline( yintercept = 0) + geom_smooth()+
  ggtitle("Residual Plot for Autism Model") + xlab("Fitted Values") + ylab("Residual Values") +theme_classic()
```

```{r, include=FALSE}
qqnorm(lm.aut$residuals)
qqline(lm.aut$residuals, col = 'red')
```

```{r, include = FALSE}
## There are some changes/
student.lmaut <- studentize(lm.aut)
leverage.lmaut <- influence(lm.aut)$hat
ggplot(data = NULL,aes(x = leverage.lmaut, y = student.lmaut)) + geom_point() + geom_hline( yintercept = 0)
```

## Concluding Thoughts

- For Typically Developing Subjects could not make distinguished correlation between motor ability and Social Deficit

- Subjects with ADHD had a slight correlation, Increased Motor skills corelated with less social deficits

- Subjects with Autism had indesernable relationship between Motor skills and Social Deficits.






